## 3. Create GeoJSON-LD
In the last chapter of this tutorial we have learned about the basic concepts of GeoJSON and GEOJSON-LD. In this chapter we will explore how to create new insightful data in GeoJSON-LD format in order to share it with users and applications. For this we will take a look at existing data in GeoJSON format and make changes to it, so that we get a sharable GeoJSON-LD file in the end. Let's get started!

```{r include=FALSE}
aasee <- toJSON(geojson_read("./data/aasee.geojson"), collapse = "\n")

featurecollection <- toJSON(geojson_read("./data/featurecollection.geojson"), collapse = "\n") 

points <- toJSON(geojson_read("./data/points.geojson"), collapse = "\n")

route <- toJSON(geojson_read("./data/route.geojson"), collapse = "\n")
```

Roland is working for a GeoIT company and he wants to collect and share geodata for touristic purposes. For that, he already collected some data in GeoJSON format and he would like to share it using the GeoJSON-LD standard. At a first step, Roland collected the area of the Aasee in Münster as a GeoJSON MultiPolygon:

```{json}
{
    "type": "Feature",
    "properties": {
        "osm_id": "13763746",
        "name": "Aasee",
        "alt_name": "",
        "basin": "retention",
        "boat": "permissive",
        "intermittent": "no",
        "maxdepth": "2",
        "name.de": "Aasee",
        "name.ru": "Аазе",
        "natural": "water",
        "salt": "no",
        "tidal": "no",
        "type": "multipolygon",
        "water": "lake",
        "website": "https://www.aaseepark.de/",
        "wikidata": "Q303434",
        "wikipedia": "de:Aasee (Münster)"
    },
    "geometry": {
        "type": "MultiPolygon",
        "coordinates": [
            [
                [
                    ...
                ]
            ]
        ]
    }
}
```

```{r echo = FALSE, message = FALSE, warning = FALSE, results = 'asis', width = 4, height = 4}
map <- leaflet(width = 1000, height = 500) %>% 
         addTiles() %>%
         setView(lng = 7.615, lat = 51.95, zoom = 14) %>%
         addGeoJSON(aasee)
map
```

Taking into account, what you have learned about the JSON-LD and GeoJSON-LD standards, think about adjustments you would have to make to the regular GeoJSON displayed above in order to transform it to the linked data format. When you are ready, click 'Next' to proceed to the next step of the tutorial.

As you have seen before, one essential part of a linked data file is the context. So, as a first step we will add the context attribute to the data:

```{json}
{
    "@context": "https://geojson.org/geojson-ld/geojson-context.jsonld", 
    "type": "Feature",
    "properties": {
        "osm_id": "13763746",
        "name": "Aasee",
        "alt_name": "",
        "basin": "retention",
        "boat": "permissive",
        "intermittent": "no",
        "maxdepth": "2",
        "name.de": "Aasee",
        "name.ru": "Аазе",
        "natural": "water",
        "salt": "no",
        "tidal": "no",
        "type": "multipolygon",
        "water": "lake",
        "website": "https://www.aaseepark.de/",
        "wikidata": "Q303434",
        "wikipedia": "de:Aasee (Münster)"
    },
    "geometry": {
        "type": "MultiPolygon",
        "coordinates": [
            [
                [
                    ...
                ]
            ]
        ]
    }
}
```
