## 3. Create GeoJSON-LD

### Introduction

In the last chapter of this tutorial we have learned about the basic concepts of GeoJSON and GEOJSON-LD. In this chapter we will explore how to create new insightful data in GeoJSON-LD format in order to share it with users and applications. For this we will take a look at existing data in GeoJSON format and make changes to it, so that we get a sharable GeoJSON-LD file in the end. Let's get started!

```{r include=FALSE}
aasee <- toJSON(geojson_read("./data/aasee.geojson"), collapse = "\n")

featurecollection <- toJSON(geojson_read("./data/featurecollection.geojson"), collapse = "\n") 

points <- toJSON(geojson_read("./data/points.geojson"), collapse = "\n")

route <- toJSON(geojson_read("./data/route.geojson"), collapse = "\n")
```

Michael wants to collect and share geodata for touristic purposes. To accomplish that, he already collected some data in GeoJSON format but he would like to share it using the GeoJSON-LD standard. As a first step, Michael collected data for two different points of interest and put it into a GeoJSON multipoint file. The first one is the Institute for Geoinformatics and the second one is the lake Aasse in Münster:

```{r echo = FALSE, message = FALSE, warning = FALSE, results = 'asis', width = 4, height = 4}
map1 <- leaflet(width = 1000, height = 500) %>% 
         addTiles() %>%
         setView(lng = 7.62, lat = 51.96, zoom = 14) %>%
         addGeoJSON(points)
map1
```
<br>
```{json}
{
    "type": "Feature",
    "geometry": {
        "type": "MultiPoint",
        "coordinates": [
            [
                7.59578,
                51.96922
            ],
            [
                7.6064,
                51.95306
            ]
        ]
    },
    "properties": {}
}
```
<br>
Taking into account, what you have learned about the JSON-LD and GeoJSON-LD standards, think about adjustments you would have to make to the regular GeoJSON displayed above in order to transform it to linked data. When you are ready, click 'Continue' to proceed to the next step of the tutorial.

### Transforming GeoJSON into GeoJSON-LD

As you might already seen in chapter 2.3 of this tutorial, one essential part of a linked data file is the context. It basically defines, how websites or applications should interpret the attributes of a linked data file. In this case, since we have GeoJSON data, we have to specify in the context attribute, how `type`, `properties` and `geometry` should be interpreted. Since there already is a publicly available context for GeoJSON data, we can just set a URL linking to the remote context file:

```{json}
{
    "@context": "https://geojson.org/geojson-ld/geojson-context.jsonld",
    "type": "Feature",
    "geometry": {
        "type": "MultiPoint",
        "coordinates": [
            [
                7.59578,
                51.96922
            ],
            [
                7.6064,
                51.95306
            ]
        ]
    },
    "properties": {}
}
```

Try to validate our adjusted GeoJSON data using the [schema.org validator](https://validator.schema.org/).

Did you encounter an error when trying to validate our adjusted data? Don't worry! This is just because schema.org does not allow for remote contexts, that are not hosted on schema.org itself. Based on tutorial chapter 2.3, think about how you could potentially resolve this problem (Hint: Have a look at the [GeoJSON context file](https://geojson.org/geojson-ld/geojson-context.jsonld)).

### GeoJSON-LD context

Instead of linking to the remote file, we could also just take the contents of the remote file and put it directly into our GeoJSON-LD file:

```{json}
{
  "@context": {
    "@version": 1.1,
    "geojson": "https://purl.org/geojson/vocab#",
    "Feature": "geojson:Feature",
    "FeatureCollection": "geojson:FeatureCollection",
    "GeometryCollection": "geojson:GeometryCollection",
    "LineString": "geojson:LineString",
    "MultiLineString": "geojson:MultiLineString",
    "MultiPoint": "geojson:MultiPoint",
    "MultiPolygon": "geojson:MultiPolygon",
    "Point": "geojson:Point",
    "Polygon": "geojson:Polygon",
    "bbox": {
      "@container": "@list",
      "@id": "geojson:bbox"
    },
    "coordinates": {
      "@container": "@list",
      "@id": "geojson:coordinates"
    },
    "features": {
      "@container": "@set",
      "@id": "geojson:features"
    },
    "geometry": "geojson:geometry",
    "id": "@id",
    "properties": "geojson:properties",
    "type": "@type"
},
    "type": "Feature",
    "geometry": {
        "type": "MultiPoint",
        "coordinates": [
            [
                7.59578,
                51.96922
            ],
            [
                7.6064,
                51.95306
            ]
        ]
    },
    "properties": {}
}
```

Try to validate our GeoJSON-LD code again and see what happens.

Amazing! We now have our first valid GeoJSON-LD file. But wait, there is more!

As you have seen before, another key feature of JSON-LD is typing. It defines how a specific JSON attribute should be interpreted by assigning an attribute type to it. "But how is this useful for sharing geographic information?" you might ask. As you might have learned, the GeoJSON standard already is very regulated in defining certain features regarding it's syntax, making it easy to interpret by applications and humans alike. It only allows specific feature types, like `Feature`, `Point` or `Multipolygon` and has to include a geometry and properties. Problems arise with the definition of properties however, since property names can be freely chosen and there are only limited data types available for properties in regular GeoJSON. This makes it difficult to work with properties in regular GeoJSON, because there is no "standardized way" of how they are defined and interpreted. To illustrate that, have a look at the example below. 

Michael also found a dataset about the lake Aasee on OpenStreetMap he would like to share. It contains a GeoJSON MultiPolygon, covering the area of the Aasee and including some useful additional information as properties, such as the name, maximum water depth and the type of water.

```{r echo = FALSE, message = FALSE, warning = FALSE, results = 'asis', width = 4, height = 4}
map2 <- leaflet(width = 1000, height = 500) %>% 
         addTiles() %>%
         setView(lng = 7.62, lat = 51.96, zoom = 13) %>%
         addGeoJSON(points) %>%
         addGeoJSON(aasee)
map2
```
<br>
```{json}
{
    "type": "Feature",
    "properties": {
        "name": "Aasee",
        "maxdepth": "2",
        "salt": "no"
    },
    "geometry": {
        "type": "MultiPolygon",
        "coordinates": [
            [
                [
                    ...
                ]
            ]
        ]
    }
}
```
<br>

As you can see, the two attributes `maxdepth` and `salt` are defined as simple string values. It would probably be more convenient though, to define the maximum depth as an actual numeric value and the water type as a boolean value. 

For example:
```{json}
{
    "type": "Feature",
    "properties": {
        "name": "Aasee",
        "maxdepth": 2,
        "salt": FALSE
    },
    ...
}
```

Things can get even more complicated if there are multiple attributes with different names and types but with the same meaning. For example you could have two different attributes for describing the water depth. One as a string and one as a numeric value:

```{json}
{
    "type": "Feature",
    "properties": {
        "name": "Aasee",
        "maxdepthmeters": 2,
        "maximumdepth": "200cm",
        "salt": FALSE
    },
    ...
}
```

Similarly, you could encounter two different attributes for describing the water type (salt or fresh water):

```{json}
{
    "type": "Feature",
    "properties": {
        "name": "Aasee",
        "maxdepth": 2,
        "isSalt": FALSE,
        "watertype": "fresh"
    },
    ...
}
```

As you can see, it can be very complicated and confusing to work with regular GeoJSON properties, because it is not transparent which properties to expect and how they are structured and to be interpreted. This is where JSON-LD typing comes in. Click on 'Continue' to learn more.

### GeoJSON-LD Typing

In this next step we are going to add LD typing to our GeoJSON properties to define how they are supposed to be interpreted. Since it makes sense to model the maximum water depth as a numeric value we are going to add the datatype `xsd:integer` to our attribute `maxdepth`:
```{json}
{
    "@context": "https://geojson.org/geojson-ld/geojson-context.jsonld",
    "type": "Feature",
    "properties": {
        "maxdepth": {
            "@type":"xsd:integer",
            "@value":"2"
        }
    },
    "geometry": {
        "type": "MultiPolygon",
        "coordinates": [
            [
                [
                    ...
                ]
            ]
        ]
    }
}
```

Think of how we could adjust our the name and type of water properties. You can use this [list of data types](https://www.w3.org/TR/xmlschema-2/#built-in-primitive-datatypes) to find out which data type might be the best fit. Click on `Continue` to see the solution.

### GeoJSON-LD Typing #2

```{json}
{
    "@context": "https://geojson.org/geojson-ld/geojson-context.jsonld",
    "type": "Feature",
    "properties": {
        "name": {
            "@type":"xsd:string",
            "@value":"Aasee"
        },
        "freshwater": {
            "@type":"xsd:boolean",
            "@value":"TRUE"
        },
        "maxdepth": {
            "@type":"xsd:integer",
            "@value":"2"
        }
    },
    "geometry": {
        "type": "MultiPolygon",
        "coordinates": [
            [
                [
                    ...
                ]
            ]
        ]
    }
}
```

### Your Turn

```{r quiz_geojsonld, echo=FALSE}
quiz(
  question("Which attribute do you atleast have to add to a GeoJSON file to make it LD? ",
    answer("@type", message = "Are you sure?"),
    answer("@context and @type", message = "The @type definition is not mandatory"),
    answer("@context", correct = TRUE, "Exactly!" ),
    allow_retry = TRUE
  ),
  question("Regarding GeoJSON, what is linked data typing especially useful for?",
    answer("to define properties", correct = TRUE),
    answer("to specify the geometry type", message = "No, the geometry of an object is already specified in the regular GeoJSON standard."),
    answer("to specify the feature type", message = "No, the feature type is already specified in the regular GeoJSON standard." ),
    allow_retry = TRUE
  )
)
```

Now it's your turn to apply the knowledge you have learned during this chapter. Michael also collected the shortest the route between the two points presented earlier as a GeoJSON line feature: 

```{r echo = FALSE, message = FALSE, warning = FALSE, results = 'asis', width = 4, height = 4}
map3 <- leaflet(width = 1000, height = 500) %>% 
         addTiles() %>%
         setView(lng = 7.62, lat = 51.96, zoom = 13) %>%
         addGeoJSON(points) %>%
         addGeoJSON(aasee) %>%
         addGeoJSON(route)
map3
```

Can you help him to transform it into a GeoJSON-LD file? You can download the data from this [link](https://raw.githubusercontent.com/LukasBaecker/geoJSON-LD_OER-module/main/data/route.geojson). Right-click on the page and save the file as a simple text file. Open the downloaded file with a text editor of your choice and try to make adjustments to it to transform it to linked data. You can use the [validator tool](https://validator.schema.org/) to check your result.
