---
title: "part2_understandingJSON"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to set up R and RStudio on your machine. We will also demonstrate
  how to install R packages from CRAN, and install the tidyverse package.
---
##   2.1. JSON

### Definition

JSON is a text-format for storing and sharing data on the web. It is developed to be understood by humans AND computers equally. The following definition is cited from [JSON.org](https://www.json.org/json-en.html).:
    
::: {#JSONdefinition .message}
JSON (JavaScript Object Notation) is a standardized lightweight data-interchange format. It is easy for humans to read and write. It is easy for machines to parse and generate. JSON is a text format that is completely language independent but uses conventions that are familiar to programmers of the C-family of languages, including `C`, `C++`, `C#`, `Java`, `JavaScript`, `Perl`, `Python`, and many others [like `R`]. These properties make JSON an ideal data-interchange language.

JSON is built on two structures:

* A collection of name/value pairs. In various languages, this is realized as an object, record, struct, dictionary, hash table, keyed list, or associative array.
* An ordered list of values. In most languages, this is realized as an array, vector, list, or sequence.

These are universal data structures. Virtually all modern programming languages support them in one form or another. It makes sense that a data format that is interchangeable with programming languages also be based on these structures.

In JSON, they take on these forms:

* An object is an unordered set of name/value pairs. 
* An object begins with `{` (left brace) and ends with `}` (right brace)
* Each name is followed by `:` (colon) and the name/value pairs are separated by `,` (comma).
:::

Taking this definition and having a look at an example we will now try to understand the concept in a more practicable way.
Have a look at the following JSON:

```json 
{
    "name": "SC Europe",
    "members":[  
        {"name":"Allan", "email":"All91cutekitties@gmail.com"},  
        {"name":"Steve", "email":"Steve.Without.N@outlook.com"},  
        {"name":"Alen", "email":"alTHEman@googlemail.com"}  
    ],
    "color": "green",
    "sports":[
        "football",
        "tennis",
        "badminton",
        "tabletennis",
        "handball",
        "basketball"
    ],
    "founded": "21-01-2005"
}  
```
As you can see, the example is easy to read and understand. Just a quick question: 

:::{.quickquestion}
What do you think might this JSON-file storing?
:::

Probably no ticket-system or a bar's menu. It is not always said that with reading the JSON-file you will understand exactly what the data is about and what it is used for, especially when the stored data gets more abstract. For example it is not very clear what is meant by color at this point. It might describe the lead color of the club but we do not know that by just looking at the JSON. Still, you mostly could get a first impression of the data by having a glimpse at the file itself.

### Setting up a JSON-file
```{r JSONpagePic2, echo=FALSE, out.width= "65%", out.extra='style="float:right; padding:10px"'}
myimg2 <- load.image("graphics/undraw_learning_sketching_nd4f.png")
grid.raster(myimg2)
```
What do we see in the example?
First, we can see, that the object is wrapped in `{` `}` as defined before. Then, we see that the object has multiple attributes that have a specific name and a value both packed in `"` `"` and divided by a `:`. The left side of the colon shows the name of the attribute while the right side shows the value. We can see, that a value can be a list (an array) of values enclosed by `[` `]` and that a value can also be an object itself.

:::{.quickquestion}
Do you see that the attribute "members" has an array of objects as value?
:::

Lets have a quick and simple task to get hands on JSON yourself:
In the following codeblock you can write the JSON file as the tasks guides you. 

:::{#luigisTask .task-message .panel}
:::{.panel-body}
Write a valid JSON file that reflects the following description: 
The JSON-file should describe a restaurant, that is called Pizza Luigi, and that stores their dishes: the first one is pizza (obviously), the second one pasta (what else) and the third one is tiramisu (delizioso!). If you are done check the solution. Additionally, there are some properties stored for the dishes: The price (a pizza costs 7€, pasta 6€ and a piece of tiramisu 10€ because it is the best in town) and it holds the information, if there is a vegan option of the dish (for pizza there is non, for pasta there is and tiramisu can be ordered vegan as well).

:::{.infomessage}
You can use the code block to insert your code but running the block will never result in a correct answer when putting in JSON-code. When finished you can compare your JSON to the one you can find by pushing the `solution` button.
:::

```{json taskLuigi2, exercise = TRUE}

```
```{json taskLuigi2-solution}
{
    "name":"Pizza Luigi",
    "dishes":[
        {"name":"pizza","price":"7","vegan_option":"false"},
        {"name":"pasta","price":"6","vegan_option":"true"},
        {"name":"tiramisu","price":"10","vegan_option":"true"}
     ]
}
```
:::
:::

### Accessing JSON-files and their values

We saved the SC Europe JSON from the top to a file, and we can load it with `R` using the `rjson` package that we installed before and loaded it in this project with `library(rjson)`. Now we can access the json file and save it to a variable.

```{r workingJSON2, include=TRUE}
sceurope <- rjson::fromJSON(file = "./data/scEurope.json")
```

Now it is your turn to access this files information. 

:::{.quickquestion}
Can you access and print ONLY Steve's email-address with the help of those two commands? Show us!
:::

The accessing of values from JSON in `R` is different than using `Javascript`, because the `rjson` parser is converting the JSON into a `R` object.

```{r accessingSCEurope2, exercise=TRUE}
#show the whole dataset
print(sceurope)
#show only when the Club was founded
print(sceurope$founded)
```
```{r accessingSCEurope2-solution}
print(sceurope$members[[2]]$email)
```

For now that is enough understanding to go on to the GeoJSON-format. In the interactive parts later on you will get more insights into working with JSON. But first let us do a short revision. If you are ready, click on the button below.
If you need any further information or want to dig deeper into the JSON-format, [JSON.org](https://www.json.org/json-en.html) is a good starting point for you.

### Revision
```{r quiz_json, echo=FALSE}
quiz(
  question("JSON stands for...",
    answer("Java Static Online Notation", message = "Sure?"),
    answer("JavaScript Object Notation", correct = TRUE, "Correct answer"),
    answer("Java Standartizised Object Notation", message = "Do you think this has something to do with Java?" ),
    allow_retry = TRUE
  ),
    question("Is it legal to have nested JSON objects?",
    answer("Yes",correct = TRUE, "As seen, an object can be the value of a property of another object"),
    answer("No", message="As seen, an object can be the value of a property of another object"),
    allow_retry = TRUE
  ),
    question("Can we dynamically access JSON-files?",
    answer("No, we can't", message = "Take a look at the tasks above, one more time."),
    answer("Yes, we can!", correct = TRUE, "And this is the awesome part about JSON!"),
    allow_retry = TRUE
  )
)
```
