---
title: "part2_understandingJSON_JSON_LD"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to set up R and RStudio on your machine. We will also demonstrate
  how to install R packages from CRAN, and install the tidyverse package.
---
##   2.3. JSON-LD

### Linked Data

Now that we worked out the idea of JSON and its GeoJSON specification we are ready to have a look at linked data. The magic is happening right now!

```{r JSONLDpicIntro, echo=FALSE, out.width= "100%", out.extra='style="float:center;"'}
myimg2 <- load.image("graphics/undraw_Inspiration_re_ivlv.png")
grid.raster(myimg2)
```
So, before getting into JSON-LD and GeoJSON-LD, we should have a look at the general concept of linked data.
Check out the video by [Manu Sporny](https://www.youtube.com/c/ManuSporny) that gives you a neat overview.

![](https://youtu.be/4x_xzT5eF5Q)

<details>
  <summary>Click to show the video below</summary>
  
  Linked Data Signature
  another Video by [Manu Sporny](https://www.youtube.com/c/ManuSporny)
![](https://youtu.be/QdUZaYeQblY)
</details>

### JSON-LD

Finally, we get to know about JSON-LD. The LD stands for 'Linked Data' (couldn't be better) and luckily we just learned what Linked Data is about. So let's jump into it.
If you would like to watch a video from [Manu Sporny](https://www.youtube.com/c/ManuSporny) explaining JSON-LD hit the play button.

![](https://youtu.be/vioCbTo3C-4)

To get a more practical view on the JSON-LD format, let's retrieve our SC EUrope example again.Here it is:

```json
{
    "name": "SC Europe",
    "members":[  
        {"name":"Allan", "email":"All91cutekitties@gmail.com"},  
        {"name":"Steve", "email":"Steve.Without.N@outlook.com"},  
        {"name":"Alen", "email":"alTHEman@googlemail.com"}  
    ],
    "color": "green",
    "sports":[
        "football",
        "tennis",
        "badminton",
        "tabletennis",
        "handball",
        "basketball"
    ],
    "founded": "21-01-2005"
}  
```
Already here, we can argue that the properties `name` could be confusing as it has different meanings for the club and its members. On the one hand we have a title of a club and on the other hand we have a human's first name. Compare the next three examples and keep that thought in mind.

:::{.quickquestion}
Can you spot the differences and think of problems that might arise?
:::

Example 1:
```json 
{
    "name": "SC Europe",
    "members":[  
        {"name":"Allan", "email":"All91cutekitties@gmail.com"},  
        {"name":"Steve", "email":"Steve.Without.N@outlook.com"},  
        {"name":"Alen", "email":"alTHEman@googlemail.com"}  
    ],
    "color": "green",
    "sports":[
        "football",
        "tennis",
        "badminton",
        "tabletennis",
        "handball",
        "basketball"
    ],
    "founded": "21-01-2005"
}  
```

Example 2:
```json 
{
    "name": "SC Europe",
    "members":[  
        {"name":"Allan", "email":"All91cutekitties@gmail.com"},  
        {"name":"Steve", "email":"Steve.Without.N@outlook.com"},  
        {"name":"Alen", "email":"alTHEman@googlemail.com"}  
    ],
    "color": "green",
    "sports":[
        "football",
        "tennis",
        "badminton",
        "tabletennis",
        "handball",
        "basketball"
    ],
    "founded": "21-01-2005"
}  
```

Example 3:
```json 
{
    "name": "SC Europe",
    "members":[  
        {"name":"Allan", "email":"All91cutekitties@gmail.com"},  
        {"name":"Steve", "email":"Steve.Without.N@outlook.com"},  
        {"name":"Alen", "email":"alTHEman@googlemail.com"}  
    ],
    "color": "green",
    "sports":[
        "football",
        "tennis",
        "badminton",
        "tabletennis",
        "handball",
        "basketball"
    ],
    "founded": "21-01-2005"
}  
```


JSON-LD Basic Markup
![](https://youtu.be/UmvWk_TQ30A)

JSON-LD Compaction and Expansion
![](https://youtu.be/Tm3fD89dqRE)




```{r todo123}
#TODO: NOCHMAL DETAILIERTER BESCHREIBEN UND SCHRITT FÜR SCHRITT ERKLÄREN WIE NE JSON-LD AUFGEBAUT IST
```

